Fusing Filters with Integer Linear Programming

### 2014-05-15 2014betterfusionforfilters/section/04-ILP.tex
Squash onto 10 pages

% -------------------------------------
\paragraph{Fusion-preventing edges} As per Megiddo\cite{megiddo1998optimal}, if there is a fusion preventing edge between two nodes we add a constraint to ensure the nodes will be placed in different clusters.
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
           \> $x_{ij}$    \> $=$   \> $1$ \>       \>       \\
           \> (for fusion-preventing edges from $i$ to $j$) 
\end{tabbing}
When combined with the precedence-preserving constraints earlier, setting $x_{ij} = 1$ also forces $\pi_i < \pi_j$.

@@

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \emptyset$)  \\
           \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\bot \in \{\iiter_{\Gamma,C}(i), \iiter_{\Gamma,C}(j)\}$)  \\
           \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \{i',j'\}$) \\
           \> \ldots
\end{tabbing}
If two nodes have different iteration sizes, they may not necessarily be fused.
If either of the nodes have no iteration size --- in which case, they are @external@ computations --- they may not be fused at all.
However, if they have parent transducers of the same size, they may be fused if both their parents are fused together, and each is fused with its respective parent.

@@

\paragraph{Array contraction}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i)            \\
           \> \ldots \\
\end{tabbing}
As per Darte's work on optimising for array contraction\cite{darte2002contraction}, we define a variable $c_i$ for each array.
An array is fully contracted if all its output edges are fused with it.
Thus, $c_i=0$ only if $\forall (i',j) \in E.\ i = i' \implies x_{ij} = 0$.
By minimising $c_i$ in the objective function, we favour solutions that reduce the number of required intermediate arrays.

@@

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> $\Sigma_{(i,j) \in E} W_{ij} x_{ij} + \Sigma_{i \in V} N c_i$  \\
           \> (if $split(i,j)$)         \\
Subject to \\
           \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j) \not\in E \wedge (j,i) \not\in E$)            \\
\\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j,\fusible) \in E$)     \\
\\
           \>             \>       \> $\pi_i < \pi_j$ \>       \>            \\
           \> (if $(i,j,\fusionpreventing) \in E$)    \\
\\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (if $split(i,j)$) \\
           \> $c_{i }$    \> $ = $ \> $ 1 $           \>       \>            \\
           \> (if $\neg(split(i,j))$) \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \emptyset$)  \\
           \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\bot \in \{\iiter_{\Gamma,C}(i), \iiter_{\Gamma,C}(j)\}$)  \\
           \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \{i',j'\} \wedge$ \\

>>>>

% -------------------------------------
\paragraph{Fusion-preventing edges} As per Megiddo~\cite{megiddo1998optimal}, if there is a fusion preventing edge between two nodes we add a constraint to ensure the nodes will be placed in different clusters.
\begin{tabbing}
MMM     \= MMM \= M \= MMM \= M \= MMM \= \kill
        \> $x_{ij}$    \> $=$   \> $1$ \>       \>       \\
        \> (for fusion-preventing edges from $i$ to $j$) 
\end{tabbing}
When combined with the precedence-preserving constraints earlier, setting $x_{ij} = 1$ also forces $\pi_i < \pi_j$.

@@

\begin{tabbing}
MMM     \= MMM \= M \= MMM \= M \= MMM \= \kill
        \> $x_{ij}$    \> $=$   \> $1$          \>       \>     \\
        \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \emptyset$
\\
\\
        \> $x_{ij}$    \> $=$   \> $1$          \>       \>     \\
        \> (if $\bot \in \{\iiter_{\Gamma,C}(i), \iiter_{\Gamma,C}(j)\}$)  \\
\\
        \> $x_{i'j}$   \> $\le$ \> $x_{ij}$     \>       \>     \\
        \> $x_{ij'}$   \> $\le$ \> $x_{ij}$     \>       \>     \\
        \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$    \>       \>     \\
        \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \{i',j'\}$)
\\
        \> \ldots
\end{tabbing}
If two nodes have different iteration sizes, they may not necessarily be fused.
If either of the nodes have no iteration size --- in which case, they are @external@ computations --- they may not be fused at all.
However, if they have parent transducers of the same size, they may be fused if both their parents are fused together, and each is fused with its respective parent.

@@

\paragraph{Array contraction}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i)            \\
           \> \ldots \\
\end{tabbing}
As per Darte's work on optimising for array contraction\cite{darte2002contraction}, we define a variable $c_i$ for each array.
An array is fully contracted if all its output edges are fused with it.
Thus, $c_i=0$ only if $\forall (i',j) \in E.\ i = i' \implies x_{ij} = 0$.
By minimising $c_i$ in the objective function, we favour solutions that reduce the number of required intermediate arrays.

@@

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> $\Sigma_{(i,j) \in E} W_{ij} x_{ij} + \Sigma_{i \in V} N c_i$  \\
           \> (if $split(i,j)$)         \\
Subject to \\
           \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j) \not\in E \wedge (j,i) \not\in E$)            \\
\\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j,\fusible) \in E$)     \\
\\
           \>             \>       \> $\pi_i < \pi_j$ \>       \>            \\
           \> (if $(i,j,\fusionpreventing) \in E$)    \\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (if $split(i,j)$) \\
           \> $c_{i }$    \> $ = $ \> $ 1 $           \>       \>            \\
           \> (if $\neg(split(i,j))$) \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \emptyset$)  \\
           \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\bot \in \{\iiter_{\Gamma,C}(i), \iiter_{\Gamma,C}(j)\}$)  \\
           \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \{i',j'\} \wedge$ \\
